\texttt{clarify}: Simulation-Based Inference for Regression Models

### 2025-09-18 rjournal_paper/Final version/clarify.tex
File management

\abstract{%
Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects and adjusted predictions. It avoids several assumptions required by the delta method that are commonly violated, especially in nonlinear models, but does not require models to be refit as when using the bootstrap. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form the sampling distribution of the quantity. \texttt{clarify} provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. \texttt{clarify} supports a large and growing number of models through its interface with the \texttt{marginaleffects} package and provides native support for multiply imputed data.
}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. The usual method for estimating the uncertainty of the derived quantities is known as the ``delta method'', which involves two approximations: 1) that the variance of the derived quantity can be represented as a first-order Taylor series, and 2) that the estimate of the derived quantity is normally distributed\footnote{In addition, most software that uses the delta method uses numerical differentiation, which also involves an approximation.}. In many cases, especially with nonlinear models, these approximation can fail badly. For example, predicted probabilities close to 0 or 1 or ratios of coefficients or probabilities typically do not have normal (or even symmetrical) distributions in finite samples, and the usual Wald-type confidence interval limits produced from delta method standard errors can include values outside the domain of the quantity of interest.

\CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating the sampling distributions of the derived quantities. Simulation-based inference is not only often more accurate than using the delta method, it is also simpler to understand and implement, as it does not require understanding Taylor series or the calculus that underlies it. This makes it more palatable to nontechnical audiences and easier to learn for students without sacrificing statistical performance.

The methodology \CRANpkg{clarify} relies on is described in King, Tomz, and Wittenberg (2000) and Rainey (2023). Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a ``posterior'' distribution, which is taken to be the sampling distribution of the derived quantity, and from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates. The delta method makes these same assumptions in addition to the approximations about the variance and the shape of the sampling distribution of the estimates of the derived quantities.

The nonparametric bootstrap is another alternative to the delta method for inference that does not require these approximations; it typically involves re-sampling individuals from the sample, fitting the model in each sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters.

More formally, we fit a regression model \(y_i = f(x_i; \beta)\), such as a linear or other generalized linear model with model coefficients \(\beta\). We assume\[
\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})
\]where \(\hat{\beta}\) is the vector of estimates of \(\beta\) and \(\Sigma_{\hat{\beta}}\) is their asymptotic covariance matrix. We define a function \(\tau(\beta)\) that represents a quantity of interest derived from the model parameters, and compute its estimate \(\widehat{\tau(\beta)}\) as \(\tau(\hat{\beta})\). Next, we take \(M\) draws \(\tilde{\beta}^{(j)}\) for \(j\in(1, \dots, M)\) from a multivariate Normal distribution with mean vector \(\mu = \hat{\beta}\) and covariance \(\Sigma = \hat{\Sigma}_{\hat{\beta}}\), where \(\hat{\Sigma}_{\hat{\beta}}\) is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of \(\tilde{\tau}=\tau(\tilde{\theta})\) as the ``posterior'' sampling distribution of \(\widehat{\tau(\beta)}\), and compute its variance as\[
\sigma^2_{\widehat{\tau(\theta)}} = \frac{1}{M}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
\]and quantile \(100(1-\alpha)\%\) confidence interval limits as \(\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]\) where \(\tilde{\tau}_{(k)}\) is the \(k\)th value of \(\tilde{\tau}\) when arranged in ascending order.

Similar functionality exists in the \pkg{CLARIFY} package in Stata\footnote{Despite the similar name, the R package \CRANpkg{clarify} and the Stata package \pkg{CLARIFY} differ in several ways, one of which is that the estimates reported by \CRANpkg{clarify} in R are those computed using the original model coefficients, whereas those reported by \pkg{CLARIFY} in Stata are those computed as the average of the simulated distribution. The R implementation avoids the ``simulation-induced bias'' described by Rainey (2023).} (Tomz, Wittenberg, and King 2003) and used to be available in the \texttt{Zelig} R package (Imai, King, and Lau 2008), though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package, which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

\hypertarget{using}{%
\subsection{\texorpdfstring{Using \CRANpkg{clarify}}{Using }}\label{using}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Fit the model to the data using modeling functions in supported packages.
\item
  Use \texttt{sim()} to take draws from the sampling distribution of the estimated model coefficients.
\item
  Use \texttt{sim\_apply()} or its wrappers \texttt{sim\_setx()}, \texttt{sim\_ame()}, and \texttt{sim\_adrf()} to compute derived quantities using each simulated set of coefficients.
\item
  Use \texttt{summary()} and \texttt{plot()} to summarize and visualize the distribution of the derived quantities and perform inference on them.
\end{enumerate}

@@

For a running example, we will use the \texttt{lalonde} dataset in the \CRANpkg{MatchIt} package (Ho et al. 2011), which contains data on 614 participants enrolled in a job training program or sampled from a survey (Dehejia and Wahba 1999). The treatment variable is \texttt{treat} and the outcome is \texttt{re78}, and all other variables are confounders. Although the original use of this dataset was to estimate the effect of \texttt{treat} on \texttt{re78}, we will use it more generally to demonstrate all of \CRANpkg{clarify}'s capabilities. In addition, we will use a transformation of the outcome variable to demonstrate applications to nonlinear models, for which the benefits of simulation-based inference are more apparent.

@@

\hypertarget{fitting-the-model}{%
\subsubsection{1. Fitting the model}\label{fitting-the-model}}

@@

\hypertarget{drawing-from-the-coefficient-distribution}{%
\subsubsection{2. Drawing from the coefficient distribution}\label{drawing-from-the-coefficient-distribution}}

@@

\begin{itemize}
\item
  \texttt{fit} -- the fitted model object, the output of the call to the fitting function (e.g., \texttt{glm()})
\item
  \texttt{n} -- the number of simulated values to draw; by default, 1000. More values will yield more replicable and precise results at the cost of speed.
\item
  \texttt{vcov} -- either the covariance matrix of the estimated coefficients, a function used to extract it from the model (e.g., \texttt{sandwich::vcovHC()} for the robust covariance matrix), or a string or formula giving a code for extracting the covariance matrix, which is passed to \texttt{marginaleffects::get\_vcov()}. If left unspecified, the default covariance matrix will be extracted from the model.
\item
  \texttt{coefs} -- either a vector of coefficients to be sampled or a function to extract them from the fitted model. If left unspecified, the default coefficients will be extracted from the model. Typically this does not need to be specified.
\item
  \texttt{dist} -- the name of the distribution from which to draw the sampled coefficients. Can be \texttt{"normal"} for a Normal distribution or \texttt{t(\#)} for a t-distribution, where \texttt{\#} represents the degrees of freedom. If left unspecified, \texttt{sim()} will decide on which distribution makes sense given the characteristics of the model (the decision is made by \texttt{insight::get\_df()} with \texttt{type\ =\ "wald"}). Typically this does not need to be specified.
\end{itemize}

@@

\texttt{sim()} uses a random number generator to draw the sampled coefficients from the sampling distribution, so a seed should be set using \texttt{set.seed()} to ensure results are replicable across sessions.

@@

\hypertarget{computing-derived-quantities}{%
\subsubsection{3. Computing derived quantities}\label{computing-derived-quantities}}

After sampling the coefficients, we will compute derived quantities on each set of sampled coefficients and store the result, which represents the ``posterior'' distribution of the derived quantity, as well as on the original coefficients, which are used as the final estimates. The core functionality is provided by \texttt{sim\_apply()}, which accepts a \texttt{clarify\_sim} object from \texttt{sim()} and a function to compute and return one or more derived quantities, then applies that function to each set of simulated coefficients. The arguments to \texttt{sim\_apply()} are below:

@@

\begin{itemize}
\item
  \texttt{sim} -- a \texttt{clarify\_sim} object; the output of a call to \texttt{sim()}.
\item
  \texttt{FUN} -- a function that takes in either a model fit object or a vector of coefficients and returns one or more derived quantities. The first argument should be named \texttt{fit} to take in a model fit object or \texttt{coefs} to take in coefficients.
\item
  \texttt{verbose} -- whether to display a progress bar.
\item
  \texttt{cl} -- an argument that controls parallel processing, which can be the number of cores to use or a cluster object resulting from \texttt{parallel::makeCluster()}.
\item
  \texttt{...} - further arguments to \texttt{FUN}.
\end{itemize}

@@

For example, we will let our derived quantity of interest be the predicted probability of the outcome for participant PSID1. We would specified our \texttt{FUN} function as follows:

@@

The resulting \texttt{clarify\_est} object contains the simulated estimates in matrix form as well as the estimate computed on the original coefficients. We will examine the sampling distribution shortly, but first we will demonstrate computing a derived quantity from the coefficients directly.

@@

\hypertarget{summarize-and-visualize-the-simulated-distribution}{%
\subsubsection{4. Summarize and visualize the simulated distribution}\label{summarize-and-visualize-the-simulated-distribution}}

@@

\texttt{plot()} displays a density plot of the resulting estimates across the simulations, with markers identifying the point estimate (computed using the original model coefficients as recommended by Rainey (2023)) and, optionally, uncertainty bounds (which function like confidence or credible interval bounds). The arguments to \texttt{plot()} are below:

@@

\begin{itemize}
\item
  \texttt{x} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be plotted if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be plotted.
\item
  \texttt{ci} -- whether to display lines at the uncertainty bounds. The default is \texttt{TRUE} to display them.
\item
  \texttt{level} -- if \texttt{ci} is \texttt{TRUE}, the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- if \texttt{ci} is \texttt{TRUE}, the method used to compute the bounds. Allowable methods include a Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using \texttt{(1-level)/2} to create a symmetric margin of error around the point estimate. The default is \texttt{"quantile"} to instead use quantile-based bounds, which are more appropriate when the distribution is non-Normal. However, quantile-based bounds may require more simulations to stabilize.
\item
  \texttt{reference} -- whether to display a normal density over the plot for each estimate. The default is \texttt{FALSE} to omit it.
\end{itemize}

@@

\begin{center}\includegraphics{figures/unnamed-chunk-8-1} \end{center}

Overlaid on the plot in red is a Normal distribution with the same mean and standard deviation as the simulated values; this is requested by setting \texttt{reference\ =\ TRUE}. From the plot, one can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling above 1 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds are clearly more appropriate than those resulting from the Normal approximation, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The plot itself is a \texttt{ggplot} object that can be modified using \texttt{ggplot2} syntax.

@@

\begin{itemize}
\item
  \texttt{object} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be displayed if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be displayed.
\item
  \texttt{level} -- the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when method (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- the method used to compute the uncertainty bounds. Allowable methods include a Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). See \texttt{plot()} above.
\item
  \texttt{null} -- an optional argument specifying the null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if \texttt{method} is \texttt{"wald"}) or an inversion of the uncertainty interval (described below). The default is not to display any p-values.
\end{itemize}

Inverting the uncertainty interval involves finding the smallest confidence level such that the null value is within the confidence bounds. The p-value for the test is one minus this level. It is only valid as a p-value when the simulated distribution of the estimates differs from its true sampling distribution under the null value by a location shift (which can be violated when the distribution of the estimates is asymmetric). However, because the p-values are invariant to monotonic transformations of the estimates and null value, as long as the distribution of \emph{some} monotonic transformation of the estimate is a location shift from its sampling distribution under the null hypothesis, the p-values will be valid.

@@

\begin{verbatim}
#>       Estimate 2.5 % 97.5 %
#> PSID1    0.976 0.891  0.996
\end{verbatim}

A benefit of quantile-based confidence bounds is that the bounds will always be in the domain of the quantity of interest, e.g., bounds for probabilities will not be above 1 or below 0 if the model always generates predicted probabilities between 0 and 1. In contrast, Wald-type symmetric confidence intervals (including those produced by the delta method) can include values outside the domain of the quantity of interest. While it is possible to compute Wald-type confidence interval bounds that respect the domain boundaries, e.g., by computing the confidence interval around the logit of the predicted probability and using the inverse logit of the bounds as the bound on the probability, quantile confidence intervals do not require the user to know which transformation of the quantity of interest will yield appropriate bounds.

@@

\begin{center}\includegraphics{figures/unnamed-chunk-10-1} \end{center}

@@

\begin{center}\includegraphics{figures/unnamed-chunk-25-1} \end{center}

>>>>

\abstract{%
Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects, average adjusted predictions, and other functions of model parameters. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form a simulated ``posterior'' distribution of the quantity from which confidence intervals can be computed. clarify provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. clarify supports a large and growing number of models through its interface with the marginaleffects package and provides native support for multiply imputed data.
}

\section{Introduction}\label{introduction}

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. Quantifying the uncertainty of these derived quantities (i.e., computing standard errors, confidence intervals, and p-values) requires additional processing. Several methods of doing so exist, including the delta method, the bootstrap, and simulation-based inference. \CRANpkg{clarify} implements simulation-based inference, which we describe below along with these other methods.

The delta method involves computing a first-order Taylor series approximation to the variance of the derived quantity, and standard Wald-based inference relies on computing quantiles based on the Normal distribution and using them to compute p-values and confidence intervals. \CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating a ``posterior'' distribution of the derived quantities. Simulation-based inference does not require understanding Taylor series or the calculus that underlies it, which can make it more palatable to nontechnical audiences and easier to learn for students without necessarily sacrificing statistical performance \citep{kingMakingMostStatistical2000, zelnerUsingSimulationInterpret2009}. Some studies have found that simulation-based inference performs as well or better than the delta method for computing derived quantities (i.e., with respect to achieving close to nominal coverage for confidence intervals), especially for complicated derived quantities and in smaller samples \citep{mackinnonConfidenceLimitsIndirect2004, holeComparisonApproachesEstimating2007, herronPostestimationUncertaintyLimited1999}. Its empirical performance has been particularly well studied in the context of mediation analysis, in which the quantities of interest are products and ratios of regression coefficients, where it has been shown to perform well relative to the delta method due to the non-Normality of these quantities \citep{tofighiMonteCarloConfidence2016, preacherAdvantagesMonteCarlo2012}.

The methodology \CRANpkg{clarify} relies on was developed by \citet{krinsky1986} and is described in \citet{kingMakingMostStatistical2000} and \citet{herronPostestimationUncertaintyLimited1999}. Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a ``posterior'' distribution, from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates, except for the approximation due to Monte Carlo error induced by sampling from a finite number of simulations (which can always be reduced by increasing the number of draws at the cost of increased computing time).

The nonparametric bootstrap is another alternative to the delta method for inference that does not require its analytic approximations \citep{efronBootstrapMethodsStandard1986}; bootstrapping typically involves re-sampling individuals from the sample, fitting the model in each bootstrap sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters. Methods for computing valid confidence intervals in cases when the quantity of interest has a complicated distribution are better developed when bootstrapping, however \citep{efronBootstrapMethodsStandard1986}.

More formally, we fit a regression model \(y_i = f(x_i; \beta)\), such as a linear or other generalized linear model with model coefficients \(\beta\). We assume \(\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})\) where \(\hat{\beta}\) is the vector of estimates of \(\beta\) and \(\Sigma_{\hat{\beta}}\) is their asymptotic covariance matrix. We define a function \(\tau(\beta)\) that represents a quantity of interest derived from the model parameters, and compute its estimate \(\widehat{\tau(\beta)}\) as \(\tau(\hat{\beta})\).

To perform simulation-based inference, we take \(M\) draws \(\tilde{\beta}^{(j)}\) for \(j\in(1, \dots, M)\) from a multivariate Normal distribution with mean vector \(\mu = \hat{\beta}\) and covariance \(\Sigma = \hat{\Sigma}_{\hat{\beta}}\), where \(\hat{\Sigma}_{\hat{\beta}}\) is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of \(\tilde{\tau}=\tau(\tilde{\beta})\) as the ``posterior'' distribution of \(\widehat{\tau(\beta)}\), and compute its variance as
\begin{equation}
\hat\sigma^2_{\widehat{\tau(\beta)}} = \frac{1}{M-1}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
\label{eq:posteriorvariance}
\end{equation}
and quantile \(100(1-\alpha)\%\) confidence interval limits as \(\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]\) where \(\tilde{\tau}_{(q)}\) is the \(q\)th value of \(\tilde{\tau}\) when arranged in ascending order (i.e., the \(q\)th quantile of the empirical cumulative distribution function of \(\tilde{\tau}\)). Simulation-based Wald-type confidence intervals can be computed as
\begin{equation}
\left[\tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{\frac{\alpha}{2}}, \tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{1-\frac{\alpha}{2}} \right]
\label{eq:siminterval}
\end{equation}
where \(Z_q\) is the \(q\)th quantile of a standard Normal distribution. The delta method-based Wald-type confidence intervals use this formula but with the first-order Taylor approximation to the asymptotic variance:
\begin{equation}
\hat\sigma^2_{\widehat{\tau(\beta)}}=\nabla\tau(\hat{\beta}) \Sigma_{\hat{\beta}} \nabla\tau'(\hat{\beta})
\label{eq:deltavariance}
\end{equation}
where \(\nabla\tau(\hat{\beta})\) is the gradient of \(\tau(\beta)\) with respect to \(\beta\) evaluated at \(\hat{\beta}\).

To compute a p-value for a hypothesis test involving the quantity of interest, i.e., \(\text{H}_0: \tau (\beta) = \tau_0\) with a given null value \(\tau_0\), we can invert the confidence interval \citep{thulinModernStatisticsWrangling2021}; that is, we find the largest value of \(\alpha\) such that \(\tau_0\) is within the confidence interval and use that \(\alpha\) as the p-value for the test. For Wald-based inference (either using the simulation-based variance or delta method-based variance), this is equivalent to performing a standard two-sided Z-test using the test statistic
\begin{equation}
Z=\frac{\tau(\hat\beta)-\tau_0}{\hat\sigma_{\widehat{\tau(\beta)}}}
\label{eq:waldz}
\end{equation}
One benefit of using the quantile p-values for inference is that equivalent tests of the same hypothesis will always yield identical p-values; for example, testing the equality of two derived quantities will yield the same p-value when comparing the difference between the quantities against a null hypothesis of 0 and the ratio of the quantities against a null hypothesis of 1, as each of these hypothesis is true if and only if the other is true.

One would expect simulation-based quantile inference, simulation-based Wald inference, and delta method-based Wald inference to align when the posterior is Normally distributed around the estimate, in which case any discrepancies would be due to Monte Carlo error in the simulated values (and therefore would shrink with increasing draws). However, for low values of \(\alpha\), it may require many draws for the simulation-based intervals to stabilize; delta method-based intervals are not subject to this error. There are a few cases in which the results might diverge: in some cases, the first-order Taylor series approximation to the variance may be poor, though in practice the approximation error is small and shrinks quickly with increasing sample size. When the posterior distribution is non-Normal but symmetric around the estimate, the quantile intervals may be more accurate (i.e., in the sense of achieving closer to nominal coverage) because they do not rely on quantiles from the Normal distribution \citep{tofighiMonteCarloConfidence2016}.

Another potential advantage quantile intervals can have over Wald-intervals is that when \emph{some} monotonic transformation of the estimate has a symmetric distribution centered around the transformed estimate, the quantile intervals can achieve correct coverage without requiring knowledge of \emph{which} transformation is required \citep{efronBootstrapMethodsStandard1986}; this is true of the quantile-based p-values as well. When the distribution is not centered around the estimate and no monotonic transformation will make it so, though, neither quantile-based nor Wald-based intervals would be expected to perform well, and quantile intervals could yield even worse coverage than Wald-based intervals, a phenomenon that occurs in the context of bootstrapping \citep{efronBootstrapMethodsStandard1986}\footnote{We thank an anonymous reviewer for pointing out a scenario in which this could occur: for a quantity of interest with a right-skewed sampling distribution, one would prefer an estimate to the right of the quantity's true value to have a confidence interval skewed to left to capture the bulk of the sampling distribution, but in practice a quantile confidence interval would also be skewed to the right. While a symmetric Wald-based interval may not have adequate coverage, the quantile-based interval could perform even worse.}. An informal falsification test for whether such a monotonic transformation exists is whether the median of the simulated estimates is aligned with the point estimate; if it is not, there is no monotonic transformation that will yield a symmetric quantile interval with the desired coverage.

\section{Related software}\label{related-software}

Similar functionality exists in the \pkg{CLARIFY} package in Stata\footnote{Despite the similar name, the R package \CRANpkg{clarify} and the Stata package \pkg{CLARIFY} differ in several ways, one of which is that the estimates reported by \CRANpkg{clarify} in R are those computed using the original model coefficients, whereas those reported by \pkg{CLARIFY} in Stata are those computed as the average of the simulated distribution. The R implementation avoids the ``simulation-induced bias'' described by \citet{raineyCarefulConsiderationCLARIFY2023}.} \citep{tomzClarifySoftwareInterpreting2003} and used to be available in the \pkg{Zelig} R package \citep{imaiCommonFrameworkStatistical2008a}, though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package \citep{marginaleffectsforthcoming}, which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

\section{\texorpdfstring{Using \CRANpkg{clarify}}{Using }}\label{using}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Fit the model to the data using modeling functions in supported packages.
\item
  Use \texttt{sim()} to take draws from the multivariate distribution of the estimated model coefficients.
\item
  Use \texttt{sim\_apply()} or its wrappers \texttt{sim\_setx()}, \texttt{sim\_ame()}, and \texttt{sim\_adrf()} to compute derived quantities using each simulated set of coefficients.
\item
  Use \texttt{summary()} and \texttt{plot()} to summarize and visualize the distribution of the derived quantities and perform inference on them.
\end{enumerate}

@@

For a running example, we will use the \texttt{lalonde} dataset in the \CRANpkg{MatchIt} package \citep{JSSv042i08}, which contains data on 614 participants enrolled in a job training program or sampled from a survey \citep{dehejiaCausalEffectsNonexperimental1999}. The treatment variable is \texttt{treat} and the outcome is \texttt{re78}, and all other variables are confounders. Although the original use of this dataset was to estimate the effect of \texttt{treat} on \texttt{re78}, we will use it more generally to demonstrate all of \CRANpkg{clarify}'s capabilities. In addition, we will use a transformation of the outcome variable to demonstrate applications to nonlinear models, for which the benefits of simulation-based inference are more apparent.

@@

\subsection{1. Fitting the model}\label{fitting-the-model}

@@

\subsection{2. Drawing from the coefficient distribution}\label{drawing-from-the-coefficient-distribution}

@@

\begin{itemize}
\item
  \texttt{fit} -- the fitted model object, the output of the call to the fitting function (e.g., \texttt{glm()})
\item
  \texttt{n} -- the number of simulated values to draw; by default, 1000. More values will yield more replicable and precise results at the cost of speed.
\item
  \texttt{vcov} -- either the covariance matrix of the estimated coefficients, a function used to extract it from the model (e.g., \texttt{sandwich::vcovHC()} for the robust covariance matrix), or a string or formula giving a code for extracting the covariance matrix, which is passed to \texttt{marginaleffects::get\_vcov()}. If left unspecified, the default covariance matrix will be extracted from the model.
\item
  \texttt{coefs} -- either a vector of coefficients to be sampled or a function to extract them from the fitted model. If left unspecified, the default coefficients will be extracted from the model. Typically this does not need to be specified.
\item
  \texttt{dist} -- the name of the distribution from which to draw the sampled coefficients. Can be \texttt{"normal"} for a Normal distribution or \texttt{t(\#)} for a T-distribution, where \texttt{\#} represents the degrees of freedom. If left unspecified, \texttt{sim()} will decide on which distribution makes sense given the characteristics of the model (the decision is made by \texttt{insight::get\_df()} with \texttt{type\ =\ "wald"}). Typically this does not need to be specified.
\end{itemize}

@@

\texttt{sim()} uses a random number generator to draw the sampled coefficients from the sampling distribution, so a seed should be set using \texttt{set.seed()} to ensure results are replicable across sessions. Using more iterations (i.e., increasing \texttt{n}) yields results that will be more stable across runs even when a seed is not set.

@@

\subsection{3. Computing derived quantities}\label{computing-derived-quantities}

After sampling the coefficients, one can compute derived quantities on each set of sampled coefficients and store the result, which represents the ``posterior'' distribution of the derived quantity, as well as on the original coefficients, which are used as the final estimates. The core functionality is provided by \texttt{sim\_apply()}, which accepts a \texttt{clarify\_sim} object from \texttt{sim()} and a function to compute and return one or more derived quantities, then applies that function to each set of simulated coefficients. The arguments to \texttt{sim\_apply()} are below:

@@

\begin{itemize}
\item
  \texttt{sim} -- a \texttt{clarify\_sim} object; the output of a call to \texttt{sim()}.
\item
  \texttt{FUN} -- a function that takes in either a model fit object or a vector of coefficients and returns one or more derived quantities. The first argument should be named \texttt{fit} to take in a model fit object or \texttt{coefs} to take in coefficients.
\item
  \texttt{verbose} -- whether to display a progress bar.
\item
  \texttt{cl} -- an argument that controls parallel processing, which can be the number of cores to use or a cluster object resulting from \texttt{parallel::makeCluster()}.
\item
  \texttt{...} -- further arguments to \texttt{FUN}.
\end{itemize}

@@

For example, we will let our derived quantity of interest be the predicted probability of the outcome for participant PSID1. We specify our \texttt{FUN} function as follows:

@@

The resulting \texttt{clarify\_est} object contains the simulated estimates in matrix form as well as the estimate computed on the original coefficients. We will examine the posterior distribution shortly, but first we will demonstrate computing a derived quantity from the coefficients directly.

@@

\subsection{4. Summarize and visualize the simulated distribution}\label{summarize-and-visualize-the-simulated-distribution}

@@

\texttt{plot()} displays a density plot of the resulting estimates across the simulations, with markers identifying the point estimate (computed using the original model coefficients as recommended by \citet{raineyCarefulConsiderationCLARIFY2023}) and, optionally, uncertainty bounds (which function like confidence or credible interval bounds). The arguments to \texttt{plot()} are below:

@@

\begin{itemize}
\item
  \texttt{x} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be plotted if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be plotted.
\item
  \texttt{ci} -- whether to display lines at the uncertainty bounds. The default is \texttt{TRUE} to display them.
\item
  \texttt{level} -- if \texttt{ci} is \texttt{TRUE}, the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- if \texttt{ci} is \texttt{TRUE}, the method used to compute the bounds. Allowable methods include the Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using \texttt{(1-level)/2} to create a symmetric margin of error around the point estimate. The default is \texttt{"quantile"} to instead use quantile-based bounds.
\item
  \texttt{reference} -- whether to display a normal density over the plot for each estimate and an indicator line for the median of the estimate. The default is \texttt{FALSE} to omit them.
\end{itemize}

@@

\begin{center}\includegraphics[alt={A plot of the density of the simulated predicted probability for participant PSID1, overlaid with a Normal distribution and a reference line for the median of the estimates}]{figures/plot1-1} \end{center}

Overlaid on the plot in red is a Normal distribution with the same mean and standard deviation as the simulated values; this is requested by setting \texttt{reference\ =\ TRUE}. From the plot, one can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling above 1 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds may be more appropriate than those resulting from the Normal approximation, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The blue reference line for the median of the estimates is close to the point estimate, suggesting it is possible for a monotonic transformation to have a symmetric distribution around the estimate\footnote{In fact, we know the inverse link function for the model (i.e., the Normal distribution function \texttt{qnorm()}) is such a transformation in this case; \CRANpkg{marginaleffects} and other packages that implement the delta method for confidence intervals around model predictions typically automatically compute confidence intervals on the transformed predictions before transforming the intervals back using the model's link function. As long as such a transformation exists, the quantile intervals will be as valid as those that rely on transforming and back-transforming.}. The plot itself is a \texttt{ggplot} object that can be modified using \texttt{ggplot2} syntax.

@@

\begin{itemize}
\item
  \texttt{object} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be displayed if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be displayed.
\item
  \texttt{level} -- the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- the method used to compute the uncertainty bounds. Allowable methods include a Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). See \texttt{plot()} above.
\item
  \texttt{null} -- an optional argument specifying the desired null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if \texttt{method} is \texttt{"wald"}) or an inversion of the uncertainty interval (described below). The default is not to display any p-values.
\end{itemize}

@@

\begin{verbatim}
#>       Estimate 2.5 % 97.5 %
#> PSID1    0.976 0.890  0.996
\end{verbatim}

@@

\begin{center}\includegraphics[alt={A plot of the density of the simulated difference between two regression coefficients, overlaid with a Normal distribution and a reference line for the median of the estimates.}]{figures/plot2-1} \end{center}

@@

\begin{center}\includegraphics[alt={A plot of the average dose-response function for the outcome against age. The line has an uncertainty band around it.}]{figures/plot8-1} \end{center}
